Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-05-28 manuscript/manuscript.Rmd
updates ms with Stan talk

All of our analyses (model fitting and simulating) were conducted in R [@Team2013]. We used the `No-U-Turn` MCMC sampler in STAN to estimate the posterior distributions of model parameters using the package 'rstan'. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 1,000 iterations after discarding an initial 1,000 iterations. We recignize such short MCMC chains may surprise some more familiar with other MCMC samplers.
We assessed convergence visually and using the @Gelman1992 diagnostic in the R package `coda' [@Plummer2006]. Scale reduction factors for all parameters were less than 1.01, indicating convergence. For the purposes of introducing stochasticity in our population models, we saved the final 1,000 iterations from each of the three MCMC chains for all parameters to be used as randomly drawn values during population simulation.

>>>>

All of our analyses (model fitting and simulating) were conducted in R [@R2013]. We used the 'No-U-Turn' MCMC sampler in Stan [@stan2014]  to estimate the posterior distributions of model parameters using the package 'rstan' [@rstan2014]. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 1,000 iterations after discarding an initial 1,000 iterations. We recignize such short MCMC chains may surprise those more familiar with other MCMC samplers (i.e. JAGS or WinBUGS), but the Stan sampler is exceptionally efficient, which reduces the number of iterations needed to achieve convergence. We assessed convergence visually and made sure scale reduction factors for all parameters were less than 1.01. For the purposes of introducing stochasticity in our population models, we saved the final 1,000 iterations from each of the three MCMC chains for all parameters to be used as randomly drawn values during population simulation. Importantly, this step alleviates the need to reduce model parameters since sampling from the full parameter space in the MCMC ensures that if a parameter broadly overlaps zero, on average the effect in the population models will also be near zero.

### 2015-05-28 manuscript/manuscript.tex
updates ms with Stan talk

Our Bayesian approach to fitting the vital rate models required choosing
appropriate priors for unknown parameters and deciding which, if any, of
those priors should be hierarchical. We decided to fit models where all
terms except climate covariates were fit by species, while the climate
covariates were fit hierarchically with species-specific coefficients
drawn from a shared `global' coefficient distribution. We did so for two
reasons: (1) the four focal species are all perennial grasses that we
expect to respond similarly to climate covariates, and (2) convergence
of climate effects at the quadrat level was much easier to achieve when
we modeled these terms hierarchically, allowing them to ``share''
statistical strength via partial pooling (Gelman and Hill 2007). So,
climate effects were modeled as:

\begin{align}
\theta_{jk} &\sim \text{Normal}(\bar{\theta_{k}}, \sigma_{k})
\end{align}

where $\bar{\theta_{k}}$ is the interspecific effect of the $k$th
climate covariate.

We used uninformative priors for all unknown parameters, specifically:

\begin{align}
\boldsymbol{\gamma, \beta, \bar{\theta}} &\sim \text{Normal}(0, 1e^{-6}) \\
\boldsymbol{\phi} &\sim \text{Normal}(0, \boldsymbol{\sigma_{\phi}}) \\
\boldsymbol{\sigma_{\phi}} &\sim e^{(\text{Gamma}(2, 0.5))} \\
\boldsymbol{\sigma_{\theta}, \sigma_{\gamma}, \sigma_{\beta}, \tau, \zeta} &\sim \text{Gamma}(0.001, 0.001)
\end{align}

All of our analyses (model fitting and simulating) were conducted in R
(Team 2013). We used the MCMC sampler in JAGS (Plummer 2003) to estimate
the posterior distributions of model parameters and the package `rjags'
to connect R to JAGS. We obtained posterior distributions for all model
parameters from three parallel MCMC chains run for 50,000 iterations,
thinned by 50, after discarding an initial 50,000 iterations. We
assessed convergence visually and using the Gelman and Rubin (1992)
diagnostic in the R package `coda' (Plummer et al. 2006). Scale
reduction factors for all parameters were less than 1.01, indicating
convergence. For the purposes of introducing stochasticity in our
population models, we saved the final 1,000 iterations from each of the
three MCMC chains for all parameters to be used as randomly drawn values
during population simulation.

@@

Gelman, A., and J. Hill. 2007. Data analysis using regression and
multilevel/hierarchical models. Page xxii, 625p.

Gelman, A., and D. B. Rubin. 1992. Inference from Iterative Simulation
Using Multiple Sequences.

@@

Plummer, M. 2003. JAGS: A Program for Analysis of Bayesian Graphical
Models Using Gibbs Sampling. Pages 20--22 \emph{in} Proceedings of the
3rd international workshop on distributed statistical computing (dSC
2003). march.

Plummer, M., N. Best, K. Cowles, and K. Vines. 2006. CODA: Convergence
Diagnosis and Output Analysis for MCMC. R News 6:7--11.

@@

@@

>>>>

Our Bayesian approach to fitting the vital rate models required choosing
appropriate priors for unknown parameters and deciding which, if any, of
those priors should be hierarchical. We decided to fit models where all
terms were fit by species. Within a species, we fit yearly size effects
and yearly intercepts hierarchically where year-specific coefficients
were drawn from global distributions representing the mean size effect
and intercept. We used flat, uninformative priors for all unknown
parameters (Appendix X).

All of our analyses (model fitting and simulating) were conducted in R
(R Core Development Team 2013). We used the `No-U-Turn' MCMC sampler in
Stan (Stan Development Team 2014a) to estimate the posterior
distributions of model parameters using the package `rstan' (Stan
Development Team 2014b). We obtained posterior distributions for all
model parameters from three parallel MCMC chains run for 1,000
iterations after discarding an initial 1,000 iterations. We recignize
such short MCMC chains may surprise those more familiar with other MCMC
samplers (i.e.~JAGS or WinBUGS), but the Stan sampler is exceptionally
efficient, which reduces the number of iterations needed to achieve
convergence. We assessed convergence visually and made sure scale
reduction factors for all parameters were less than 1.01. For the
purposes of introducing stochasticity in our population models, we saved
the final 1,000 iterations from each of the three MCMC chains for all
parameters to be used as randomly drawn values during population
simulation. Importantly, this step alleviates the need to reduce model
parameters since sampling from the full parameter space in the MCMC
ensures that if a parameter broadly overlaps zero, on average the effect
in the population models will also be near zero.

@@

@@

@@

R Core Development Team. 2013. R: A language and environment for
statistical computing.

@@

Stan Development Team. 2014a. Stan: A C++ Library for Probability and
Sampling, Version 2.5.0.

Stan Development Team. 2014b. Rstan: the R interface to Stan, Version
2.5.0.
